\documentclass{beamer}

\mode<presentation>

\title{Write You A Garbage Collector: The Basics}
\author{Matthew Russo}
\institute{Reinventing the Wheel}
\date{February 12th, 2020}

\begin{document}
 
\frame{\titlepage}

\begin{frame}
  \frametitle{Who Am I?}
  \begin{enumerate}
    \item Consistently reinventing the wheel
    \item Enjoy listening to and discovering new music. Would like to one day make music!
    \item Software Engineer for AWS working on ECR
    \item Knowledge sharing around these more `mysterious' parts of software development.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Who Are You?}
  \begin{enumerate}
    \item Name & Pronouns
    \item Why you joined this group?
    \item Any specific topics of interest? Projects you've worked on?
    \item Anything else you'd like us to know
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Why are we here? (and what are we not here for)}
\end{frame}

\begin{frame}
  \frametitle{Terminology}
  \begin{enumerate}
    \item Stack & Heap
    \item Runtime
    \item Allocator
    \item Collector
    \item Mutator
    \item Root
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Four basic garbage collection algorithm families}
  \begin{enumerate}
    \item Mark-Sweep (Tracing)
    \item Mark-Compact (Tracing)
    \item Copying (Tracing)
    \item Reference Counting
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Marking in Tracing collectors}
  \begin{enumerate}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Mark-Sweep Collection}
\end{frame}

\begin{frame}
  \frametitle{Mark-Compact Collection}
\end{frame}

\begin{frame}
  \frametitle{Copying Collection}
\end{frame}

\begin{frame}
  \frametitle{Reference Counting Collection}
\end{frame}


\end{document}
